---
output: html_document
---
### Reproducible Research Course Assignment 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###### Load library
```{r  echo = TRUE}
library(ggplot2)
library(dplyr)
```

###### Read the activity file and store in activity variable
```{r   echo = TRUE}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```

###### Part1
###### Calculate the mean total of steps in a day
```{r   echo = TRUE}
totStepsPerDay <- aggregate(steps~date, data = activity, sum,
                            na.rm = TRUE)
str(totStepsPerDay)
```

###### Plot the histogram for total mean steps in a day
```{r echo = TRUE}
hist(totStepsPerDay$steps, xlab = "Total steps", 
                           ylab ="Number of Days",
                            main = "Mean total steps per day")
```


###### Calculate the mean of total steps in a day
```{r echo = TRUE}
meanOfSteps <- mean(totStepsPerDay$steps)
meanOfSteps
```


###### Calculate the median of total steps in a day
```{r  echo = TRUE}
medianOfSteps <- median(totStepsPerDay$steps)
medianOfSteps
```

###### Part 2
###### Calculate average number of steps taken, averaged across all days

```{r   echo = TRUE}
meanStepsPerInterval <- aggregate(steps~interval, data = activity, mean, na.rm = TRUE)
str(meanStepsPerInterval)
```

###### Plotting average number of steps taken, averaged across all days

```{r   echo = TRUE}
with(meanStepsPerInterval,  plot(interval,steps, type = "l")) 
```

###### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r  echo = TRUE}
maxInterval <- meanStepsPerInterval[which.max(meanStepsPerInterval$steps),]$interval
maxInterval
```
 
###### Part 3
###### Calculate the mean per intervals across all days and store in meanStepsPerIterval
###### Replace the NA values with meanPerInterval and store the entire set as activity2
###### Group steps by date for both activity and activity2
```{r echo = TRUE}

meanStepsPerIterval <- aggregate(steps ~ interval, activity, mean)
activity2 <- transform(activity, steps = ifelse(is.na(activity$steps), 
                    meanStepsPerIterval$steps[match(activity$interval, 
                    meanStepsPerIterval$interval)], activity$steps))
stepsPerDayWithNa = aggregate(steps~date, activity, sum)
stepsPerDayWithoughtNa = aggregate(steps~date, activity2, sum)

summary(stepsPerDayWithoughtNa)
```

###### Potting the histogram of steps per day after replacing NA's 
```{r echo = TRUE}
hist(stepsPerDayWithNa$steps, xlab = "Steps", ylab = "Frequency", col = "orange",
    main = paste("Frequency of steps per day after replacing NA values"))
```

###### To see the difference, will plot two more graphs
###### First one will plot "Sum of steps per day withought NA" (g1)
###### Second one will plot "Sum of steps per day with NA" (g2)
###### If we examine two graphs, we can see empty spaces in g2 due to NA's



###### Plot the graph g1 : Sum of steps per day withought NA

```{r   echo = TRUE}
g1 <- ggplot(data = stepsPerDayWithoughtNa, aes(date, steps))
g1 <- g1 + geom_bar(stat = "identity", color = "orange", width = .1)
g1 <- g1 + labs(xlab = "Date", ylab = "Steps")
g1 <- g1 + labs(title = "Sum of steps per day withought NA")
print(g1)
```

###### Plot the graph g2 : Sum of steps per day with NA
```{r  echo = TRUE}
g2 <- ggplot(data = stepsPerDayWithNa, aes(date, steps))
g2 <- g2 + geom_bar(stat = "identity", color = "green", width = .1)
g2 <- g2 + labs(xlab = "Date", ylab = "Steps")
g2 <- g2 + labs(title = "Sum of steps per day with NA")
print(g2)
```


###### Part 4
###### Create a new factor variable for storing a day as "weekday" or "weekend"
````{r echo = TRUE}
activity2 <- activity
activity2 <- mutate(activity2,  day = ifelse( weekdays(activity2$date) == "Saturday"|
                                 weekdays(activity2$date) == "Sunday",
                                 "weekend",
                                 "weekday"))  
````

###### Create a time series plot for 5 minute interval and average number of steps taken on weekend or weekday
```{r echo = TRUE}

meanWeekDayEnd <- aggregate(steps~day+interval, activity2, mean)
g <- ggplot(data = meanWeekDayEnd, aes(interval,steps, color = day))
g <- g + geom_line() + facet_grid(day~.)
print(g)

```

